---
title: "Dawid-Skene-anesthesia"
output: pdf_document
---

```{r setup, include=FALSE}
library("rater")
require(rjags)
require(coda)
require(here)
require(rstan)
options(mc.cores = parallel::detectCores())
rstan_options(auto_write = TRUE)

NT <- 2 # number of trials
set.seed(123)
```

```{r compile, include=FALSE}
rater(anesthesia, dawid_skene())
```


```{r jags and stan setup, include=FALSE}
data("anesthesia", package = "rater")
load(file = here("data","anesthesia.rda"))
N <- nrow(anesthesia) # total number of annotations
K <- 4 # number of annotation categories
I <- 45 # number of items
J <- 5 # number of annotators
ii <- anesthesia$item # item index for annotation n
jj <- anesthesia$rater # annotator for annotation n
y <- anesthesia$rating # annotation for observation n


# alpha <- rep(3,K)
# n <- 8 # pseudocount of hypothetical prior observations
# p <- 0.6 # approximate probability of a correct rating
# beta <- array((n*(1-p)/(K-1)),c(J,K,K))
# beta
# for(j in 1:J){
#   for (k in 1:K){
#     beta[j,k,k] <- n*p
#   }
# }
  

data_jags <- list(N=N, y=y, K=K, I=I, J=J, ii=ii, jj=jj)
# data_stan <- list(N=N, y=y, K=K, I=I, J=J, ii=ii, jj=jj,beta=beta, alpha=alpha)

iterations <- 6000
burnin <- floor(iterations/2)
chains <- 3

parameters = c("pi", "theta","z")
```

```{r sampling, include=FALSE}
t <- 1 # trial number

repeat{
  # record jags computation time 
  assign(paste("time_jags", t, sep = ""), 
         system.time(model.fit <- 
                       jags.model(file =here("Models","Dawid-Skene.txt"), 
                                  data=data_jags, 
                                  n.chains=chains))["elapsed"]
         + system.time(model.samples <- 
                         coda.samples(model.fit, parameters, 
                                      n.iter=iterations))["elapsed"]) 
  
  # record jags ess
  assign(paste("ess_jags", t, sep = ""), effectiveSize(model.samples)[0:84])
  
  # record jags time per min ess 
  assign(paste("timeperminess_jags", t, sep = ""), 
         eval(as.name(paste("time_jags", t, sep = "")))/
           min(eval(as.name(paste("ess_jags", t, sep = "")))))
  
  # record jags time per mann ess 
  assign(paste("timepermeaness_jags", t, sep = ""), 
         eval(as.name(paste("time_jags", t, sep = "")))/
           mean(eval(as.name(paste("ess_jags", t, sep = "")))))
  
  # record jags model
  assign(paste("model_jags", t, sep = ""), model.samples)
  
  # record stan computation time
  assign(paste("time_stan", t, sep = ""),
         system.time(model.fit <- rater(anesthesia, dawid_skene()))["elapsed"])

  model_fit <- get_stanfit(model.fit)
  
  # assign(paste("time_stan", t, sep = ""), 
  #        system.time(model_fit <- 
  #                      suppressMessages(
  #                        stan(file = here("Models","DawidSkene.stan"), 
  #                             data = data_stan, iter=iterations, 
  #                             chain=chains, warmup=burnin)))["elapsed"])
  
    
  # record stan ess
  assign(paste("ess_stan", t, sep=""), summary(model_fit)$summary[,"n_eff"][0:84])
  
  # record stan time per min ess 
  assign(paste("timeperminess_stan", t, sep = ""), 
         eval(as.name(paste("time_stan", t, sep = "")))/
           min(eval(as.name(paste("ess_stan", t, sep = "")))))
 
  # record stan time per mawn ess 
  assign(paste("timepermeaness_stan", t, sep = ""), 
         eval(as.name(paste("time_stan", t, sep = "")))/
           mean(eval(as.name(paste("ess_stan", t, sep = ""))))) 
  
  
  # record stan model
  assign(paste("model_stan", t, sep = ""), model_fit)
 
  # record rater model
  assign(paste("model_rater", t, sep = ""), model.fit) 
  
  # record jags marginalised computation time 
  assign(paste("time_jags_mar", t, sep = ""), 
         system.time(model.fit <- 
                       jags.model(file =here("Models","Dawid-Skene-marginalised.txt"), 
                                  data=data_jags, 
                                  n.chains=chains))["elapsed"]
         + system.time(model.samples <- 
                         coda.samples(model.fit, parameters, 
                                      n.iter=iterations))["elapsed"]) 
  
  # record jags marginalised ess
  assign(paste("ess_jags_mar", t, sep = ""), effectiveSize(model.samples)[0:84])
  
  # record jags marginalised time per min ess 
  assign(paste("timeperminess_jags_mar", t, sep = ""), 
         eval(as.name(paste("time_jags_mar", t, sep = "")))/
           min(eval(as.name(paste("ess_jags_mar", t, sep = "")))))
  
  # record jags marginalised time per mann ess 
  assign(paste("timepermeaness_jags_mar", t, sep = ""), 
         eval(as.name(paste("time_jags_mar", t, sep = "")))/
           mean(eval(as.name(paste("ess_jags_mar", t, sep = "")))))
  
  # record jags marginalised model
  assign(paste("model_jags_mar", t, sep = ""), model.samples)
  if (t == NT) break
  
  t <- t+1
}

```

```{r gathering result, include=FALSE}
trial <- 1:NT

# vector of result of the trials

temp <- paste("timeperminess_jags", trial, sep = "")
lis_timeperminess_jags <- sapply(temp, as.name)
lis_timeperminess_jags <- sapply(lis_timeperminess_jags, eval)

temp <- paste("timeperminess_stan", trial, sep = "")
lis_timeperminess_stan <- sapply(temp, as.name)
lis_timeperminess_stan <- sapply(lis_timeperminess_stan, eval)

temp <- paste("timeperminess_jags_mar", trial, sep = "")
lis_timeperminess_jags_mar <- sapply(temp, as.name)
lis_timeperminess_jags_mar <- sapply(lis_timeperminess_jags_mar, eval)

temp <- paste("timepermeaness_jags", trial, sep = "")
lis_timepermeaness_jags <- sapply(temp, as.name)
lis_timepermeaness_jags <- sapply(lis_timepermeaness_jags, eval)

temp <- paste("timepermeaness_stan", trial, sep = "")
lis_timepermeaness_stan <- sapply(temp, as.name)
lis_timepermeaness_stan <- sapply(lis_timepermeaness_stan, eval)

temp <- paste("timepermeaness_jags_mar", trial, sep = "")
lis_timepermeaness_jags_mar <- sapply(temp, as.name)
lis_timepermeaness_jags_mar <- sapply(lis_timepermeaness_jags_mar, eval)
```

\newpage
## Result
### Main results using the minimum and the mean ESS across parameter

```{r boxplot1,  echo=FALSE, fig.height=4, fig.width=5.5}

boxplot(lis_timeperminess_jags, lis_timeperminess_stan,lis_timeperminess_jags_mar,main="time per effective sample(min)", names = c("jags", "stan", "marginalised"))

```

```{r boxplot2, fig.height=4, fig.width=5.5, echo=FALSE}
boxplot(lis_timepermeaness_jags, lis_timepermeaness_stan,lis_timepermeaness_jags_mar,main="time per effective sample(mean)", names = c("jags", "stan", "marginalised"))
```

\newpage
### Number of Effective Samples per second
Comparison between JAGS and Stan
```{r ess plots, echo=FALSE, fig.align='center'}
order <- sort(names(ess_jags1))
for(i in 1:NT){ 
  jags_norm_ess <- eval(as.name(paste("ess_jags", i, sep = "")))[order][1:84]/
                   eval(as.name(paste("time_jags", i, sep = "")))
              
  stan_norm_ess <-  eval(as.name(paste("ess_stan", i, sep = "")))[order][1:84]/
                   eval(as.name(paste("time_stan", i, sep = "")))
  
  print(plot(jags_norm_ess,stan_norm_ess, xlab="JAGS", ylab="Stan", main = "Number of Effective Samples per second"))
  
}

```

## Convergence Diagnostic

\newpage
### Traceplot for Stan
```{r traceplot stan, echo=FALSE}
stan_p <- length(names(model_stan1)) # number of continuous parameter
for(i in 1:NT){
  for(j in seq(from = 1, to = stan_p, by =10)){
    sprintf("Trial %d", i)
    print(traceplot(eval(as.name(paste("model_stan", i, sep = ""))), 
                    par = names(eval(as.name(paste("model_stan", i, sep = ""))))[j:(j+9)]))
  }
}
```

\newpage
### Traceplot for JAGS

```{r traceplot discrete, echo=FALSE}
for(i in 1:NT){
  sprintf("Trial %d", i)
  par(mfrow = c(3,3))
  coda::traceplot(eval(as.name(paste("model_jags", i, sep = ""))))
}

```

\newpage

### Looking at Rhat
Looking at continuous paramters only
For JAGS we used the gelman rubin diagnostic and looked at the upper CI

```{r rhat for stan and discrete model, echo=FALSE}
for(i in 1:NT){
  print(paste("Trial",i))
  rhat_stan <- mcmc_diagnostics(eval(as.name(paste("model_rater", i, sep = ""))))[,"Rhat"]
  disc <- gelman.diag(eval(as.name(paste("model_jags", i, sep = ""))), multivariate = FALSE)
  rhat_jags <- disc$psrf[1:84,"Upper C.I."][order[1:84]] # 129 in total, first 84 continuous
  print(paste("max Rhat for stan:",max(rhat_stan)))
  print(paste("max Rhat for JAGS:",max(rhat_jags)))
  plot(rhat_jags, rhat_stan, xlab="Rhat JAGS", ylab="Rhat Stan")
}

```

\newpage
## Distribution of latent class
### For Stan
```{r error matrices stan, echo=FALSE}
for (i in 1:NT){
  p <- plot(eval(as.name(paste("model_rater", i, sep = ""))), "raters")
  plot(p)
}

```


\newpage
### For JAGS

```{r function for plotting table , include=FALSE}
# taken fromhttps://elliotnoma.wordpress.com/2015/07/18/how-to-plot-a-table-of-values-in-r/
plot_table <- function(d, marginColor,main="", text.cex=1.0)
{
plot(c(-1,ncol(d)),c(0,nrow(d)+1), type="n", xaxt="n", yaxt="n", xlab="",ylab="", bty="n")
title(main, line = -10)
for (c in 1:ncol(d)) {
#rect(c-1, nrow(d), c, nrow(d) + 1, col=marginColor)
text(c-.5,nrow(d) +.5,colnames(d)[c], cex=text.cex)
}

for (r in 1:nrow(d)) {
#rect(-1, r-1, 0, r, col=marginColor)
text(-.5, r-.5,rownames(d)[nrow(d) - r + 1], cex=text.cex)
}

for (r in 1:nrow(d))
for (c in 1:ncol(d)) {
rect(c-1, r-1, c, r)
text(c-.5,r-.5,d[nrow(d) - r + 1,c], cex=text.cex)
}
}
```


```{r error matrices discrete, echo=FALSE, fig.height=3.5, fig.width=3.5}
## error matrices for discrete model

for(i in 1:NT){
  sprintf("Trial %d", i)
  disc <- summary(eval(as.name(paste("model_jags", i, sep = ""))))
  thetas <- disc$statistics[order,"Mean"][5:84]
  start <- 1
  print(paste("JAGS trial",i, ":distribution of latent classes"))
  while(start < length(thetas)){
   tot_entry <- 16
   accuracy <- thetas[start:(start+tot_entry-1)]
   error_matrix <- matrix(accuracy, nrow = 4, byrow=TRUE)
   #print(error_matrix)
   plot_table(round(error_matrix,2), "gray", 
              main= paste("rater",(start-1)/16+1,sep=""), text.cex=1)
   start <- start + tot_entry
  }
  writeLines('\n')
}

```